---
title: "R Tutorial"
author: "Mattingly"
date: "2/10/2020"
output: pdf_document
---

getwd()
setwd("/Users/petermattingly/Desktop/")

### creating a notebook chunk
'control' + 'option', then 'i'

```{r}

```

### running individual lines of code
'command' then 'return'


### assignment operator <-


### creating pipe operator %>%
'command' 'shift' 'm' = 


### libraries and packages

```{r}
install.packages('data.table')
install.packages('tidyverse')
library(data.table)
library(tidyverse)
```



### import data

```{r}
# 50 flowers from 3 species of iris
# measurements (cm) for sepal and petal width and length
data(iris)
head(iris)
```

```{r}
dim(iris)
glimpse(iris)
```


### missing values

```{r}
table(is.na(iris))
```


```{r}
iris$Sepal.Length[which(iris$Sepal.Length == -1 )] <- NA
```



```{r}
summary(iris, na.rm = TRUE)
```

```{r}
levels(iris$Species)
```

```{r}
levels(iris$Sepal.Length)
```

```{r}
class(iris$Sepal.Length)
```


### create row id variable

```{r}
### base r
iris_base <- cbind(Flowers_base = rownames(iris), Iris)
glimpse(iris_base)
```

```{r}
### dplyr
iris <- iris %>% mutate(Flowers = row_number())
glimpse(iris)
```


### subsetting

##### subsetting by value

```{r}
### base r
setosa <- iris[iris$Species == "setosa",]
glimpse(setosa)
```

```{r}
### dplyr
setosa_tidy <- iris %>% filter(Species = "setosa")
glimpse(setosa_tidy)
```



##### subsetting by columns

```{r}
### base r
iris_length <- iris[, c(1,3,5,9)]
glimpse(iris_length)
```


```{r}
### dplyr
iris_length_dplyr <- iris %>% dplyr::select(matches("(Length|Species)"))
glimpse(iris_length_dplyr)
```


```{r}
### base r
iris_width <- iris[, c(2,4,5)]

### dplyr
iris_width_dplyr <- iris %>% dplyr::select(matches("(Width Species)"))
glimpse(iris_width_dplyr)
```



### frequencies

```{r}
### range
range(iris$Sepal.Length)
### frequency table
table(iris$SepalLength)
```

```{r}
### histogram
hist(iris$Sepal.Length)
### boxplot
boxplot(iris$Sepal.Length, horizontal == TRUE)
```


```{r}
quantile(iris$Sepal.Length)
```


```{r}
quantile(iris$Sepal.Length, probs = c(0.05, 0.5, 0.95))
```




```{r}
iris$Levels <- cut(iris$Sepal.Length, 3, include.lowest = TRUE, 
                   labels = c("low","med","high","very high"))
table(iris$Levels, iris$Sepal.Length)
glimpse(iris)
```


```{r}
iris_stats <- iris %>%
  group_by(Species, Levels) %>%
  summarize(Freq = n())
head(iris_stats)
```

```{r}
iris_stats <- iris %>%
  group_by(Species, Levels) %>%
  summarize(Freq = n(),
            Mean = mean(Sepal.Length),
            Median = median(Sepal.Length),
            Stdv = sd(Sepal.Length)) %>%
head(iris_stats)
```


### reshape iris data

```{r}
iris_reshape <- iris %>%
  ### convert data from 'wide' to long'
  ### designate 'key' variable for grouping all remaining variables
  ### designate 'value' variable to take on all corresponding values
  gather(key = "key", value = "value", -Species, -Flowers) 
head(iris_reshape)
```


```{r}
iris_reshape <- iris_reshape %>%
  ### split up 'key' using "."
  ### designate the names for the two new variables (e.g., 'part', 'measure')
  separate(key, c("Part","Measure"), sep =  "\\.")
head(iris_reshape)
```


```{r}
iris_reshape <- iris_reshape %>%
  ### convert data from 'long' to wide'
  ### take nominal levels of 'measure' as new variables
  spread(Measure, value)
head(iris_reshape)
```


### reshape iris data

```{r}
iris_reshape <- iris %>%
  gather(key = "key", value = "value", -Species, -Flowers) %>%
  separate(key, c("Part","Measure"), sep =  "\\.") %>%
  spread(Measure, value)

dim(iris_reshape)
glimpse(iris_reshape)
```


### save and open iris data

```{r}
write.csv(iris_reshape, "~/Desktop/r tutorial/iris_reshape.csv", row.names = FALSE)
```

```{r}
iris_reshape <- fread("~/Desktop/r tutorial/iris_reshape.csv")
glimpse(iris_reshape)
```



### new data

```{r}
data(nhtemp)
str(nhtemp)
```



```{r}
nhtemp <- as.data.frame(nhtemp)
nottem <- nottem %>% rename(Temp = x)
names(nhtemp)[1] <- "Temp"
```


### google: r convert time series to dataframe

```{r}
nhtemp <- data.frame(temp = as.matrix(nhtemp), year = time(nhtemp))
glimpse(nhtemp)
```

```{r}

```





