---
title: "Untitled"
author: "Mattingly"
date: "2/13/2020"
output: pdf_document
---

getwd()
setwd("/Users/petermattingly/Desktop/")


## Tutorials:
##  which type of plot works best
##  https://www.data-to-viz.com/

##  colors
##  https://www.datanovia.com/en/blog/ggplot-colors-best-tricks-you-will-love/

##  overall aesthetics
##  http://urbanspatialanalysis.com/dataviz-tutorial-mapping-san-francisco-home-prices-using-r/


### creating a notebook chunk
'control' + 'option', then 'i'

```{r}

```


### classic packages

```{r}
install.packages('data.table')
install.packages('tidyverse')
```


### packages - colors

```{r}
install.packages("crayon")
install.packages("prismatic")
install.packages("magrittr")
install.packages("scales")
devtools::install_github("EmilHvitfeldt/prismatic")
```

### libraries

```{r}
## previous libraries
library(data.table)
library(tidyverse)

## new libraries
library(crayon)
library(prismatic)
library(magrittr)
library(scales)
```



### import data

```{r}
data(iris)
glimpse(iris)
``` 


### create row 'id' variable named

```{r}
iris <- iris %>% mutate(ID = row_number())
glimpse(iris)
```


### reshape iris data - no notes

```{r}
iris_reshape <- iris %>%
  gather(key = "key", value = "value", -Species, -ID) %>%
  separate(key, c("Part","Measure"), sep =  "\\.") %>%
  spread(Measure, value)
```



### dimensions / glimpse

```{r}
dim(iris_reshape)
glimpse(iris_reshape)
```


### reshape iris data - w/ notes

```{r}
iris_reshape <- iris %>%
  ### convert data from 'wide' to long'
  ### designate 'key' variable for converting the remaining variables into a single categorical variable
  ### designate 'value' variable to take on all corresponding values of those categories
  gather(key = "key", value = "value", -Species, -ID) 
glimpse(iris_reshape)
```

```{r}
iris_reshape <- iris_reshape %>%
  ### split up 'key' using "."
  ### designate the names for the two new variables (e.g., 'part', 'measure')
  separate(key, c("Part","Measure"))
glimpse(iris_reshape)
```

```{r}
iris_reshape <- iris_reshape %>%
  ### convert data from 'long' to wide'
  ### take nominal levels of 'measure' as new variables
  spread(Measure, value)
glimpse(iris_reshape)
```


```{r}
setosa <- iris_reshape %>% filter(Species == "setosa")
glimpse(setosa)
```


### ggplot

####  "grammar of graphics"; graph componenets as building blocks
####  flexibility; rigor

### create base

```{r}
ggplot(data = iris_reshape)
```



### scatter/point plot

```{r}
ggplot(iris_reshape) +
  ## geometry point
  ## aesthetics - "something you can see"; a mapping between variable and a visual cue
  geom_point(aes(x = Length,
                 x = Width))
```



### scatter/point plot

```{r}
ggplot(iris_reshape) +
  geom_point(y = Length,
             x = Width)
```



### scatter/point plot

```{r}
ggplot(iris_reshape) +
  geom_point(aes(y = Length,
                 x = Width))
```



### aesthetic mapping
####  shapes

```{r}
# squares
### 0 - blank square
### 15 - one-color square
### 22 - two-color square

# circles
### 1 - blank circle
### 16 - one-color circle
### 21 - two-color circle

# triangle
### 2 - blank triangle
### 17 - one-color triangle
### 24 - two-color triangle
```


### aesthetic mapping - shapes

```{r}
ggplot(iris_reshape) +
  geom_point(aes(y = Length,
                 x = Width, 
                 shape = Species))
```



### aesthetic mapping - shapes

```{r}
levels(iris_reshape$Species)
```



### aesthetic mapping - shape manual

```{r}
ggplot(iris_reshape) +
  geom_point(aes(y = Length,
                 x = Width, 
                 shape = Species,
                 color = Species)) + 
  scale_shape_manual(name = "Species",
                     labels = c("Setosa","Versicolor","Virginica"),
                     values = c(22,21,24))
```



### aesthetic mapping - colors
### color brewer: http://colorbrewer2.org/

```{r}
ggplot(iris_reshape) +
  geom_point(aes(y = Length,
                 x = Width, 
                 color = Species,
                 shape = Species)) + 
  scale_shape_manual(name = "Species",
                     labels = c("Setosa","Versicolor","Virginica"),
                     values = c(22,21,24)) +
    scale_color_manual(name = "Species",
                     labels = c("Setosa","Versicolor","Virginica"),
                     values = c("#9ecae1","#4292c6","#08519c"))
```


### color palettes 

```{r}
blue <- colorRampPalette(c("blue", "white"))
blue(20)
plot(rep(1, 20), col = blue(20), pch = 19, cex = 15)
```




####  see more info on prismatic package
###   https://www.hvitfeldt.me/blog/manipulating-colors-with-prismatic/

```{r}
color(plasma(10))
```



```{r}
color(viridis_pal()(5))
```


```{r}
show_col(viridis_pal()(5))
```




### colors

```{r}
####  create palette
palette1 <- color(terrain.colors(10))

####  colors
palette1 %>% plot()

####  colors + label
scales::show_col(palette1)


####  colors + label
scales::show_col(clr_tritan(palette1))

clr_tritan(palette1) %>% plot()


clr_tritan(palette1)
```

### colors - tritan

```{r}
#clr_saturate(palette1, 0.5) %>% plot()
#clr_desaturate(palette1, .50) %>% plot()
#clr_negate(palette1) %>% plot()
new_pal <- clr_protan(palette1) %>% plot()
new_pal <- clr_rotate(palette1, 80)
clr_tritan(new_pal) %>% plot()
```



### aesthetic mapping - labels and colors

```{r}
ggplot(iris_reshape) +
  geom_point(aes(y = Length,
                 x = Width, 
                 color = Width,
                 fill = Species, 
                 shape = Species)) + 
      ### scale_color_gradient automatically goes from light to dark
  #scale_color_gradient(low = "red", high = "blue") +
      ### scale_color_gradient2 automatically goes from dark to light
      ### scale_color_gradientn enables you to create a custom sscale
  scale_color_gradientn(colors = (viridis_pal()(5))) +
  labs(title = "Title",
       subtitle = "Subtitle",
       y = "Y - Axis",
       x = "X - Axis",
       caption = "Caption")
```


### aesthetic mapping - scales

```{r}
ggplot(iris_reshape) +
  geom_point(aes(y = Length,
                 x = Width, 
                 color = Width,
                 fill = Species, 
                 shape = Species)) + 
      ### scale_color_gradient automatically goes from light to dark
  #scale_color_gradient(low = "red", high = "blue") +
      ### scale_color_gradient2 automatically goes from dark to light
      ### scale_color_gradientn enables you to create a custom sscale
  scale_color_gradientn(colors = (viridis_pal()(5))) +
  labs(title = "Title",
       subtitle = "Subtitle",
       y = "Y - Axis",
       x = "X - Axis",
       caption = "Caption") +
  scale_y_continuous(limits = c(3,8,)) +
  scale_x_continuous(limits = c(0.9,4.1)) 
```



### reduce redundancy in manual blocks

```{r}
### labels
lab <- c("Setosa", "Versicolor", "Virginica")

### colors
cols <- c("setosa" = "#006FA6FF", "versicolor" = "#00E64DFF", "virginica" == "#440154FF")

### shapes
shapes <- c(22, 21, 24)
```



### aesthetic mapping - colors

```{r}
graph1 <- ggplot(iris_reshape) +
  geom_point(aes(y = Length,
                 x = Width, 
                 color = Species,
                 fill = Species, 
                 shape = Species,
                 size = Length), alpha = .5) + 
  
  scale_color_manual(name = "Speciez",
                     labels = lab,
                     values = cols) +
  scale_fill_manual(labels = lab,
                     values = cols) +
  scale_shape_manual(labels = lab,
                     values = shapes) +
graph1
ggsave("~/graph1.png", graph1, width = 16, height = 9, device = "png")
```



### aesthetic mapping - colors

```{r}
graph1 <- ggplot(iris_reshape) +
  geom_point(aes(y = Length,
                 x = Width, 
                 color = Species,
                 fill = Species, 
                 shape = Species,
                 size = Length), alpha = .5) + 
  
  scale_color_manual(name = "Species",
                     labels = lab,
                     values = cols) +
  scale_fill_manual(labels = lab,
                     values = cols) +
  scale_shape_manual(labels = lab,
                     values = shapes) +
  scale_size(guide = FALSE) +
  
  
  stat_smooth(aes(y = Length,
                 x = Width), method = "lm", alpha = 0.65, size = 1.5, se = FALSE) +
  
  labs(title = "Title",
       subtitle = "Subtitle",
       y = "Y - Axis",
       x = "X - Axis",
       caption = "Caption") +
  
  facet_wrap(~Species)
graph1
ggsave("~/graph1.png", graph1, width = 16, height = 9, device = "png")
```


### flip coordinates

```{r}
graph1 + coord_flip()
```



### count bar plot

```{r}
ggplot(iris_reshape) +
  geom_bar(aes(Species))
```



### count bar plot - fill

```{r}
ggplot(iris_reshape) +
  geom_bar(aes(Species, 
               fill = Part))
```



### count bar plot - fill; reverse

```{r}
ggplot(iris_reshape) +
  geom_bar(aes(Species, 
               fill = Part),
           position = position_stack(reverse = TRUE))
```




### mean bar plot

```{r}
ggplot(iris_reshape) +
  geom_col(aes(x = Species, 
               y = Width))
```



### create median dataframe

```{r}
iris_median <- iris_reshape %>%
  group_by(Species, Part) %>%
  summarize(Length = median(Length),
            Width = median(Width))
glimpse(iris_median)
```



### group medians

```{r}
ggplot(iris_reshape) +
  geom_bar(aes(y = Width, fill = Part, x = Species),
           position = "dodge", stat = "identity", alpha = .85) +
  
  labs(title = "Title",
       subtitle = "Subtitle",
       y = "Y - Axis",
       x = "X - Axis",
       caption = "Caption")
```



### group medians - colors / adjust y axis

```{r}
ggplot(iris_median) +
  geom_bar(aes(y = Width, fill = Part, x = Species),
           position = "dodge", stat = "identity") +
  
  scale_fill_manual(name = "Part",
                    labels = c("Petal","Sepal"),
                    values = c("#FDE725FF","#3B528BFF")) +
  
    labs(title = "Title",
       subtitle = "Subtitle",
       y = "Y - Axis",
       x = "X - Axis",
       caption = "Caption") +
  
  coord_cartesian(ylim = c(1.8, 4))
```



### median bar plot - scales


```{r}
graph2 <- ggplot(iris_median) +
  geom_bar(aes(y = Width, fill = Part, x = Species),
           position = "dodge", stat = "identity") +
  
  scale_fill_manual(name = "Part",
                    labels = c("Petal","Sepal"),
                    values = c("#FDE725FF","#3B528BFF")) +
  
  scale_y_continuous(labels = comma) +
  
  scale_x_discrete(labels = toupper(levels(iris_median$Species))) + 
  
  labs(title = "Title",
       subtitle = "Subtitle",
       y = "Y - Axis",
       x = "X - Axis",
       caption = "Caption")
graph2
```



### themes

```{r}
graph2 + theme_clean()
ggsave("~/Desktop/graph2.png", graph2, width = 11.5, height = 9, device = "png")
```




```{r}
plotTheme <- function(base_size = 12) {
  theme(
    text = element_text( color = "black"),
    plot.title = element_text(size = 18,colour = "black"),
    plot.subtitle = element_text(face="italic"),
    plot.caption = element_text(hjust=0),
    axis.ticks = element_blank(),
    
    panel.background = element_blank(),
    panel.grid.major = element_line("grey80", size = 0.1),
    panel.grid.minor = element_blank(),
    
    strip.background = element_rect(fill = "grey80", color = "white"),
    strip.text = element_text(size=12),
    
    axis.title = element_text(size=8),
    axis.text = element_text(size=8),
    axis.title.x = element_text(hjust=1),
    axis.title.y = element_text(hjust=1),
    plot.background = element_blank(),
    legend.background = element_blank(),
    legend.title = element_text(colour = "black", face = "italic"),
    legend.text = element_text(colour = "black", face = "italic"))
}
```



```{r}
graph2 + plotTheme()
```














































